Parsing Postfix log files

### 2007-04-18 doc/logparser.tex
Added Policy server section. Added more XXX items, including placeholder log coverage section.

% Better typesetting of URLs.
\usepackage{url}
% Include images
\usepackage[final]{graphicx}
\renewcommand{\labelenumii}{\roman{enumii}:}

@@

XXX WHAT IS THE CONCLUSION???

@@

Postfix is a highly configurable, high performance, secure and scalable
Mail Transport Agent.  It features extensive anti-spam restrictions,
allowing a mail administrator to deploy those restrictions which they judge
suitable for their needs, rather than a fixed set chosen by Postfix's
author.  These restrictions can be selectively applied, combined and
bypassed on a per-client, per-recipient or per-sender basis, allowing
varying levels of defense and or permissiveness.  Postfix leverages simple
lookup tables to support arbitrarily complicated user-defined sequences of
restrictions and exceptions, with the ultimate in flexibility being the
facility to consult an external process which can implement whatever logic
is required: some users can be restricted to sending mail on the third
Tuesday after pay day only, if that is of any use.\footnote{This example
may actually be useful in a payroll system.  More commonly encountered
scenarios are restricting posting to large mailing lists (e.g.\ sending
weekly offers to customers) to out of office hours, rate limiting clients,
or checking SPF records (see http://www.openspf.org/ or
http://en.wikipedia.org/wiki/Sender\_Policy\_Framework for details)}
Administrators can also supply their own rejection messages to make it
clear to senders why exactly their mail was rejected.  Unfortunately this
flexibility has a cost: complexity in the logs generated.  While it is easy
to use \texttt{grep(1)} to determine the fate of an individual email,
following the journey an email takes through Postfix can be quite
difficult.  The logs tend to follow a 90\%-10\% pattern: 90\% of the time
the journey is simple, but the other 10\% of the time requires 90\% of the
code.\footnote{These numbers don't have a solid scientific basis, they're
based on gut feeling from writing and debugging the software.}

@@

@@

@@

@@

XXX CARL SUGGESTS EXPANDING ON THIS SOME MORE, BUT I DON'T KNOW HOW\@.
PERHAPS WHEN THE ALGORITHM IS PROPERLY EXPLAINED IT'LL BE EASIER\@?  MAKE
REFERENCES TO THE ALGORITHM, EXPLAIN MORE CLEARLY HOW THE FIELDS ARE USED,
HOW THE RULE MATCHES, ETC\@.

@@

@@

    \item Each new Postfix release requires new rules to be written to cope
        with the new messages.  Similarly new RBLs, new policy servers and
        new administrator defined rejection messages require new rules.

>>>>

% Better typesetting of URLs.
\usepackage{url}
% Include images
\usepackage[final]{graphicx}
% Change how nested enumerate environments are labelled.
\renewcommand{\labelenumii}{\roman{enumii}:}

@@

XXX WHAT IS THE CONCLUSION???  FIRST THING IS TO COMPARE TO OTHER PARSERS
REVIEWED

@@

Postfix is a highly configurable, high performance, secure and scalable
Mail Transport Agent.  It features extensive anti-spam restrictions,
allowing a mail administrator to deploy those restrictions which they judge
suitable for their needs, rather than a fixed set chosen by Postfix's
author.  These restrictions can be selectively applied, combined and
bypassed on a per-client, per-recipient or per-sender basis, allowing
varying levels of defense and or permissiveness.  Postfix leverages simple
lookup tables to support arbitrarily complicated user-defined sequences of
restrictions and exceptions, with policy servers as the ultimate in
flexibility.  Administrators can also supply their own rejection messages
to make it clear to senders why exactly their mail was rejected.
Unfortunately this flexibility has a cost: complexity in the logs
generated.  While it is easy to use \texttt{grep(1)} to determine the fate
of an individual email, following the journey an email takes through
Postfix can be quite difficult.  The logs tend to follow a 90\%-10\%
pattern: 90\% of the time the journey is simple, but the other 10\% of the
time requires 90\% of the code.\footnote{These numbers don't have a solid
scientific basis, they're based on gut feeling from writing and debugging
the software.}

A policy server is an external program that accepts state information from
Postfix for each SMTP command not rejected by an earlier restriction; the
policy server can utilise that state information to implement whatever
logic is required.  E.g.\ some users can be restricted to sending mail on
the third Tuesday after pay day only --- this example may actually be
useful in a payroll system to prevent spam or (worse) phishing mails with
faked sender addresses.  More commonly encountered scenarios are:

    \item Checking Sender Policy Framework records.  SPF records specify
        which mail servers are allowed to send mail claiming to be from a
        particular domain.  The intention is to reduce faked sender
        addresses, which will in turn reduce spam, backscatter and
        joe-jobs; however there has been a lot of resistance to the
        proposal because it breaks or vastly complicates some features of
        SMTP.  For more details see \url{http://www.openspf.org/} or
        \url{http://en.wikipedia.org/wiki/Sender_Policy_Framework} XXX ADD
        REFERENCES FOR SPAM BACKSCATTER AND JOE-JOBS TO THE APPENDIX.

    \item Greylisting, a technique that temporarily rejects mail where the
        triple of (sender, recipient, remote IP address) is unknown; the
        assumption is that it's uneconomical for a spammer to retry, but
        that a legitimate mail server must retry.  Sadly spammers are using
        increasingly complex and well written programs to distribute spam,
        frequently using an ISP provided SMTP server from a compromised
        machine on its network, so greylisting is slowly becoming less
        useful, but it does block a large percentage of spam mail at the
        moment.  See \url{http://www.greylisting.org/} for full details.

    \item Using a scoring system such as Policyd-weight where 
        tests accumulate points against the sending system --- if the
        eventual score is too high the mail is rejected.
        \url{http://www.policyd-weight.org/}

    \item Policyd performs the following: \textit{Greylisting,
        Sender-(envelope, SASL or host / ip)-based throttling (on messages
        and/or volume per defined time unit), Recipient rate limiting,
        Spamtrap monitoring / blacklisting, HELO auto blacklisting and HELO
        randomization preventation.\/}  \url{http://policyd.sourceforge.net}

Example attributes taken from \newline
\url{http://www.postfix.org/SMTPD_POLICY_README.html}:

\begin{tabular}[]{ll}

    request                 & smtpd\_access\_policy     \\
    protocol\_state         & RCPT                      \\
    protocol\_name          & SMTP                      \\
    helo\_name              & some.domain.tld           \\
    queue\_id               & 8045F2AB23                \\
    sender                  & foo@bar.tld               \\
    recipient               & bar@foo.tld               \\
    recipient\_count        & 0                         \\
    client\_address         & 1.2.3.4                   \\
    client\_name            & another.domain.tld        \\
    reverse\_client\_name   & another.domain.tld        \\
    instance                & 123.456.7                 \\

@@

XXX REFER TO THE GRAMMAR COMPARISON SECTION ONCE ITS WRITTEN.

@@

XXX COMPARE AND CONTRAST WITH CONTEXT FREE GRAMMARS, OTHER GRAMMARS.

XXX DESCRIBE IT IN LEFT: RIGHT PRODUCTION TERMS, IF POSSIBLE.

@@

        XXX RERUN AGAIN ONCE POSTFIX 2.3 RULES HAVE BEEN ADDED TO SEE WHAT
        THE DIFFERENCE IS.  THERE ARE 110 POSTFIX 2.2 RULES AS OF
        2007/04/18

@@

XXX CARL SUGGESTS EXPANDING ON THIS SOME MORE, BUT I DON'T KNOW HOW\@.
PERHAPS WHEN THE ALGORITHM IS PROPERLY EXPLAINED IT'LL BE EASIER\@?  MAKE
REFERENCES TO THE ALGORITHM, EXPLAIN MORE CLEARLY HOW THE FIELDS ARE USED,
HOW THE RULE MATCHES, AND SO ON\@.

@@

\section{Parsing coverage}

XXX THIS SECTION NEEDS TO BE WRITTEN

\label{parsing-coverage}

\subsection{Log lines covered}

\label{log-lines-covered}

\subsection{Mails covered}

\label{mails-covered}

@@

    \item Each new Postfix release requires new rules to be written to cope
        with the new log lines.  Similarly new RBLs, new policy servers and
        new administrator defined rejection messages require new rules.
